---
title: My dynamic report
output: html_fragment
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)

```

```{r}
library(tidyverse)


#import Data from google sheets 
logged_data <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vReF3vvFTYztjwdvP619aEGeKT4gSAO3CgVGqMzVhwYX-cG1owc7YRPli22YF7K8-uo9QDpu_4cFxUd/pub?gid=743329457&single=true&output=csv")

#Renamed columns of imported data 
latest_data <- logged_data %>% rename(Bus_Stop_Arrival_Time = 2,
  AT_Mobile_Wait_Time = 3,
  Bus_Scheduled_Arrival_Time = 4,
  Departure_Bus_Stop_Number = 5, 
  Boariding_Bus_Number = 6,
  Day_Of_Week = 7,
  Time_Boarded_Bus = 8,
  AT_Mobile_Occupancy_Status= 9,
  Bus_Type = 10,
  Double_Empty_Seats_Available = 11,
  Sitting_Or_Standing = 12,
  Time_Off_Bus = 13,
  Destination_Bus_stop_number = 14,
  Weather = 15
)

#Creating data columns that are data-time type objects intead of strings for calculation 
#The functions as.Date() and as.POSIXct() I have learnt by googling and found on this website
#https://www.neonscience.org/resources/learning-hub/tutorials/dc-convert-date-time-posix-r
#I have also discussed the usage of these functions with Liza 


#Generate new column Date using TimeStamp column date from chr to date 
latest_data$Date <- as.Date(
  latest_data$Timestamp,
  format = "%m/%d/%Y %H:%M:%S"
)


# Convert Bus_Stop_Arrival_Time data type chr to date-time 
latest_data$Bus_Stop_Arrival_TS <- as.POSIXct(
  paste(latest_data$Date, latest_data$Bus_Stop_Arrival_Time),
  format = "%Y-%m-%d %H:%M:%S"
)

# Convert Time_Boarded_Bus data type chr to date-time 
latest_data$Time_Boarded_Bus_TS <- as.POSIXct(
  paste(latest_data$Date, latest_data$Time_Boarded_Bus),
  format = "%Y-%m-%d %H:%M:%S"
)

# Convert Time_Off_Bus data type chr to date-time 
latest_data$Time_Off_Bus_TS <- as.POSIXct(
  paste(latest_data$Date, latest_data$Time_Off_Bus),
  format = "%Y-%m-%d %H:%M:%S"
)

# Generate new column Trip duration calculated  from Time_Off_Bus_TS - Time_Boarded_Bus_TS
latest_data$Trip_Duration <- latest_data$Time_Off_Bus_TS - latest_data$Time_Boarded_Bus_TS


# Generate new column actual_bus_wait_time from Time_Boarded_Bus_TS - Bus_Stop_Arrival_TS
latest_data$Actual_Wait_Time <- latest_data$Time_Boarded_Bus_TS - latest_data$Bus_Stop_Arrival_TS

#Convert Actual_Wait_Time from seconds to minutes 
latest_data$Actual_Wait_Minutes <- as.numeric(latest_data$Actual_Wait_Time, units = "mins")


```

```{css}

```

# Summary of Bus Wait Times

**Nature of the data:**\
This section uses **quantitative, time-based data**. The variables `Actual Wait Minutes` and `At Mobile Wait Time` record numeric wait times in minutes for each bus trip. The data is observational and consists of repeated measurements collected at multiple timestamps for different trips.

**What I am analysing:**\
I am analysing differences between the **actual time waited at the bus stop** and the **predicted wait time from the AT Mobile app**. Calculating the **minimum, maximum, and average wait times** allows identification of trends in bus service reliability and the consistency of app predictions.

The minimum time I actually waited for a bus to come was `r min(latest_data$Actual_Wait_Minutes, na.rm = TRUE)` minutes, the maximum was `r max(latest_data$Actual_Wait_Minutes, na.rm = TRUE)` minutes, and the average wait time was `r round(mean(latest_data$Actual_Wait_Minutes, na.rm = TRUE), 2)` minutes.

The minimum wait time for a bus recorded on AT Mobile was `r min(latest_data$AT_Mobile_Wait_Time, na.rm = TRUE)` minutes, the maximum was `r max(latest_data$AT_Mobile_Wait_Time, na.rm = TRUE)` minutes, and the average wait time was `r round(mean(latest_data$AT_Mobile_Wait_Time, na.rm = TRUE), 2)` minutes.

# Bar charts

## Actual Wait Minutes vs At MobileWait Time Bar Chart

**Nature of the data:**\
This plot visualises **quantitative time-based variables** (`Actual Wait Minutes` and `AT Mobile Wait Time`) for each trip. Each timestamp represents a repeated observation of a single bus journey.

**What I am analysing:**\
The chart compares **actual vs predicted wait times** to determine whether buses arrive earlier or later than expected, and to evaluate the accuracy and consistency of the app’s predictions over multiple trips.

```{r}


# Actual_Wait_Minutes vs AT_Mobile_Wait_Time
ggplot(latest_data, aes(x = Timestamp)) +
  geom_col(aes(y = Actual_Wait_Minutes), 
           fill = "blue", width = 0.4) +
  geom_col(aes(y = AT_Mobile_Wait_Time), 
           fill = "red", width = 0.4, 
           position = position_nudge(x = 0.2)) +
  labs(title = "Actual Wait Minutes vs AT Mobile Wait Time",
       subtitle =  "blue = Actual Wait Minutes, red = At Mobile Wait Time",
       caption = "Source: My Obsevation log",
       x = "Date", 
       y = "Wait Time (minutes)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

## Bus Occupancy by Day of the Week

**Nature of the data:**\
This section uses **categorical data**. `AT Mobile Occupancy Status` indicates crowding levels on the bus, while `Day Of Week` identifies the day each observation was made. The data is observational and collected over multiple trips.

**What I am analysing:**\
I am analysing how **bus occupancy varies by day of the week**, identifying patterns in crowding. This shows which days tend to have more standing passengers or more available seats, highlighting trends in passenger load across the week.

```{r}
#AT_Mobile_Occupancy_Status vs Day_Of_Week
latest_data %>%
  ggplot() +
  geom_bar(aes(y = AT_Mobile_Occupancy_Status, 
               fill = Day_Of_Week)) +
  labs(title = "AT Mobile Occupancy Status vs Weather",
       caption = "Source: My Obsevation log",
       x = "AT Mobile Occupancy Status Count",
       y = "AT Mobile Occupancy Status type ")
```

## Bus Occupancy by Weather

**Nature of the data:**\
This section also uses **categorical data**. `AT Mobile Occupancy Status` shows crowding levels, while `Weather` records conditions during each trip. Data is repeated over multiple trips and weather types.

**What I am analysing:**\
I am analysing how **bus occupancy changes under different weather conditions**, to understand whether weather affects crowding and passenger comfort. The plot helps reveal trends such as more standing passengers on rainy days or more available seating on sunny days.

```{r}
#AT_Mobile_Occupancy_Status vs Weather",
latest_data %>%
  ggplot() +
  geom_bar(aes(y = AT_Mobile_Occupancy_Status, fill = Weather)) +
  labs(title = "AT Mobile Occupancy Status vs Weather",
       caption = "Source: My Obsevation log",
       x = "AT Mobile Occupancy Status Count",
       y = "AT Mobile Occupancy Status type ")

```
